---
title: "Marsupial shape and phylogenetics"
author: "Leonie Lange-Hodgson"
date: "22 January 2019"
output: html_document
editor_options: 
  chunk_output_type: inline
---


# load required libraries

#I will be using the following R packages in these analyses; geomorph, ape, plyr, abind, and geiger.

```{r}

# Note: need to have dispRity ticked in the library before you can install landVR

library(rgl)
library(RRPP)
library(geomorph)
library(ape)
library(plyr)
library(abind)
library(geiger)
library(caper)
library(devtools)
library(phytools)
library(nlme)
library(dispRity)
library(landvR)
library(png)
library(viridis)
library(stringr)
library(scales)

Dviv <- read.ply("../Data/Dasyurus_viverrinus_ascii.ply")

load (file = "../Data/Processed/processed_data_for_mars_phylo_shape.rda")

```




#Images for partitions


```{r}
#colouring partition spheres 

 #provides the numbers of the parts
  PartLevels = unique(part.gp[])
  #colour scheme ;-)
  Colours <- c("hotpink", "purple")
  
  #Mesh material
  
 
  #Make mean shape that the points can be plotted on
  Ref=mshape(GPA_Allmean$coords)
  
  #Or, for 3d plotting, call up the coordinates of the mean specimen; remember that the radius of the spheres might need to increase because this is not GPA'ed
  Ref <- Data[,,7]
  
  Part=list()
  ##subset the landmarks according to the partitions
  for(i in 1:length(PartLevels)){
    Part[[i]]<-which (part.gp[] == PartLevels[[i]])
  }
  
  ##colours the spheres for each partition on the mean shape
  
  open3d()
  
  for (i in 1:length(PartLevels)){
    spheres3d( Ref[(Part[[i]]),]*1000, col=Colours[i], lit=TRUE, radius = 900, asp=F)
    
  }

  shade3d (Dviv, color="grey")
  
```


#Evolutionary allometry

```{r}
#Geomorph update 3.1.0 on the 27th March removed procD.Allometry so using the function procD.lm instead. 
Phyall<-procD.pgls(coords~Csize, tree_synch, data=GPA_Allmean, iter=999)
summary(Phyall)


Phyallbasi<-procD.pgls(coords~Csize, tree_synch, data=GPA_Basimean, iter=999)
summary(Phyallbasi)

plotAllometry(Phyallbasi, size=GPA_Basimean$Csize, method = "RegScore")

Phyallrest<-procD.pgls(coords~Csize, tree_synch, data=GPA_Restmean, iter=999)
summary(Phyallrest)


```


#PlotRefto target - Function plots shape differences between a reference and target specimen
#vector: a plot showing the vector displacements between corresponding landmarks in the reference and target specimen is shown


# Visualize main differences OF WHAT AND WHY? YOU MIGHT ALSO ADD SOME DESCRIPTION OF THE RESULTS HERE, WHICH YOU CAN LATER PUT INTO RESULTS SECTION.
```{r}

#Below is a starting point for customizing your PlotRefToTarget plots. I recommend playing with this! This then gets fed into plotRefToTarget as per below.
PRTT=gridPar(pt.bg="green", pt.size = 1)

plotRefToTarget(PCA_mean$pc.shapes$PC1min, PCA_mean$pc.shapes$PC1max, method="vector", mag=0.6,gridPars = PRTT)
plotRefToTarget(PCA_mean$pc.shapes$PC2min, PCA_mean$pc.shapes$PC2max, method="vector") #Back to standard grey
```



Allometry in geometric morphometrics refers to size-related changes of morphological traits. 
e.g. does brain size increase as body size increases?
looking at overall shape however

#Csize has an effect on shape of the landmarks

##When analysing the allometry of the remaining landmark on the skull the results showed the effect size of Csize is quite small however there is a fairly significant p-value suggesting that Csize does have an effect on shape

```{r}
#Geomorph update 3.1.0 on the 27th March removed procD.Allometry so using the function procD.lm instead. 
Allom<-procD.lm(coords~Csize, logsz=TRUE, data=GPA_Allmean, iter=10000, print.progress=FALSE)
summary(Allom)


AllBasi<-procD.lm(coords~Csize, data=GPA_Basimean, iter=1000, print.progress=FALSE)
summary(AllBasi)

plotAllometry(AllBasi, size=GPA_Basimean$Csize, method = "RegScore")

AllRest<-procD.lm(coords~Csize, logsz=TRUE, data=GPA_Restmean, iter=10000, print.progress=FALSE)
summary(AllRest)

```



#Compare evolutionary rates of partitions and phylogenetic signal



Evolutionary rates

#Phylogenetic signal


```{r}
#### SHAPE RATE ANALYSES
### Compare evolutionary rates between partitions
# Calculate rates for each partition


parevo.rate <- 
  compare.multi.evol.rates(GPA_Allmean$coords,  part.gp, tree_synch,iter = 1000)
summary(parevo.rate)

# Does phylogenetic signal differ? do for the other two as well and csize

physignal(GPA_Allmean$coords, tree_synch, iter = 1000)

physignal(GPA_Basimean$coords, tree_synch, iter = 1000)

physignal(GPA_Restmean$coords, tree_synch, iter = 1000)

```

#PGLS for Diet and Locomotion (look at all three GPAs)


```{r}
#Combine the GPA geomorph dataframe with the two columns I want from the speclist data frame

#Do this for each of the GPA dataframes


Allandlist<-mutate(GPA_Allmean, diet=speclist$Diet, locomotion=speclist$Locomotion)

Basiandlist<-mutate(GPA_Basimean, diet=speclist$Diet, locomotion=speclist$Locomotion)

Restandlist<-mutate(GPA_Restmean, diet=speclist$Diet, locomotion=speclist$Locomotion)


#First look at diet and locomotion for all landmarks


Alldiet_inter<-procD.pgls(coords ~ diet, tree_synch, data = Allandlist, iter = 999)
summary(Alldiet_inter)

AllLoco_add<-procD.pgls(coords ~ locomotion, tree_synch, data = Allandlist, iter=999)
summary(AllLoco_add)


# Next look at basicranium landmarks

Basidiet_inter<-procD.pgls(coords~diet, tree_synch, data = Basiandlist, iter = 999)
summary(Basidiet_inter)

BasiLoco_inter<-procD.pgls(coords~locomotion, tree_synch, data = Basiandlist, iter=999)
summary(BasiLoco_inter)

plot3d(Basiandlist$coords[,,8], asp=TRUE); text3d(Basiandlist$coords[,,8], texts=c(35:58))

                                            
                                                   
#Then at the rest of the skull landmarks

RestDiet_inter<-procD.pgls(coords~diet, tree_synch, data = Restandlist, iter = 999)
summary(RestDiet_inter)

RestLoco_add<-procD.pgls(coords~locomotion, tree_synch, data = Restandlist, iter=999)
summary(RestLoco_add)


```


#To understand the main variation in shape between marsupial species I run a Principal Component Analysis (PCA) on the coordinate data. Using the function plotTangentSpace from the geomorph package, the resulting graph will provide the two PC axes that explain the most variation between species coordinates. To understand the variance of each PC axis, I will run the function pc.sumary$importance to view the eigenvalues. These eigenvalues aid in further analyses.

#Each PC represents a particular aspect of the overall shape variation, see more in the chunk

```{r}

#Run PCA for all three partitions
PCA_mean <- plotTangentSpace(GPA_Allmean$coords, label=dimnames(GPA_Allmean$coords)[[3]])

PCA_basi <- plotTangentSpace(GPA_Basimean$coords, label=dimnames(GPA_Basimean$coords)[[3]])

PCA_rest <- plotTangentSpace(GPA_Restmean$coords, label=dimnames(GPA_Restmean$coords)[[3]])

```

Setting up some plotting parameters

```{r}

colclade<-c("mediumpurple2", "darkorange","yellowgreen","dodgerblue3", "firebrick1", "violetred", "forestgreen", "darkorchid", "darkblue") #Nine specific colours for the nine clades
names(colclade) <- levels(speclist$Clade) # assign levels to colours
col.gp <- colclade[match(speclist$Clade, names(colclade))] # magic!
head(colclade) # see the result

#Making a short form vector of species names of type for plotting
Gen.sp <- matrix(unlist(strsplit(as.character(speclist$Specimen), "_")),
ncol=2, byrow = T)
# string split is used to separate a character string
head(Gen.sp)
Gen.sp <- paste(substr(Gen.sp[,1], 1, 1), substr(Gen.sp[,2], 1, 3), sep=".")
head(Gen.sp) # see the result

```


#Phylogenetic signal in PC scores; the below is just a prototype and not working very well. I don't understand the need to ultrametricise for OU estimates and I am unsure how reliable this all is given my data.


```{r}

PCA_score_list <- list("Basi"=PCA_basi$pc.scores, "Rest"=PCA_rest$pc.scores, "All"=PCA_mean$pc.scores)


##Make function to determine W to later feed into loop

fit.W <- function(variable, tree){


fitBM <-fitContinuous(tree,variable, model="BM")
fitOU <-fitContinuous(tree,variable, model="OU")
fitEB <-fitContinuous(tree,variable, model="EB")

AICs <- c(fitBM$opt$aicc,fitOU$opt$aicc,fitEB$opt$aicc)
AICmin <- AICs-min(AICs)
W <- exp(-0.5*AICmin)/sum(exp(-0.5*AICmin))

W <- setNames(W, c("BM", "OU","EB"))

return(W)

}

#Now run W determinant on all

W_list <- list()

#number of partition PCAs (3 - basi, rest and mean)

for (Partitions in 1:3) {
  
        W_list[[Partitions]] <- fit.W (PCA_score_list[[Partitions]][,1], compute.brlen(tree_synch))
        }


W_list

#The below should not be run on non-BM data but here it is for now.

PCA_score_list <- list("Basi"=PCA_basi$pc.scores, "Rest"=PCA_rest$pc.scores, "All"=PCA_mean$pc.scores)
                       
                       
phylosig_comp <- matrix(nrow=3, ncol = 4)
rownames (phylosig_comp)<- names(PCA_score_list)
colnames(phylosig_comp) <- dimnames(PCA_score_list[[1]])[[2]][1:4]

#number of PCAs (3 - basi, rest and eman)
for (PCAs in 1:3){
      #number of PCs considered (4 here but doesn't matter)
      for (PCscores in 1:4){
        
        phylosig_comp [PCAs,PCscores] <- phylosig(tree_synch,PCA_score_list[[PCAs]][,PCscores], method="K", nsim=1000)
        
      }
  }

as.data.frame(phylosig_comp)



```

PCA_score_list[[3]][,4]

phylosig(tree_synch,PCA_score_list[1]$pc.scores[,2],method="K", nsim=1000)

phylosig(tree_synch, PCA_basi$pc.scores[,1],method="K", nsim=1000)
phylosig(tree_synch, PCA_basi$pc.scores[,2],method="K", nsim=1000)
phylosig(tree_synch, PCA_basi$pc.scores[,3],method="K", nsim=1000)
phylosig(tree_synch, PCA_basi$pc.scores[,4],method="K", nsim=1000)
phylosig(tree_synch, PCA_basi$pc.scores[,5],method="K", nsim=1000)

phylosig(tree_synch, PCA_basi$pc.scores[,1],method="K", nsim=1000)
phylosig(tree_synch, PCA_basi$pc.scores[,2],method="K", nsim=1000)
phylosig(tree_synch, PCA_basi$pc.scores[,3],method="K", nsim=1000)
phylosig(tree_synch, PCA_basi$pc.scores[,4],method="K", nsim=1000)
phylosig(tree_synch, PCA_basi$pc.scores[,5],method="K", nsim=1000)


#Plotting the PCA

```{r}

plot.new()
par(mfrow=c(2,1))
par(mar=c(1.2,5,1.2,1.2))

plot(PCA_score_list$Basi[,1], PCA_score_list$Basi[,2], type="p",pch=21, bg=col.gp, cex=1.5, ylab="Basicranium");

plot(PCA_score_list$Rest[,1], PCA_score_list$Rest[,2], type="p",pch=21, bg=col.gp, cex=1.5, ylab="Rest of Skull")

plot(PCA_score_list$All[,1], PCA_score_list$All[,2], type="p",pch=21, bg=col.gp, cex=1.5)


legend(x=10.9,y=-0.125,
legend=c("Caenolestid","Dasyurom.", "Didelphid","Macropod",  expression(italic("Dromiciops")), "Peramelem.","Petaroid", "Phalangeroid",
"Vombatiform"),
pt.bg=c(unique(aggregate(col.gp ~ speclist$Clade, FUN=unique)[,-1]),"light grey"),
pch=c(rep(21,11),24), cex=1, pt.cex=1, ncol = 3, bg= "light grey", box.col="light grey",
text.width = 0.52, y.intersp = 1.3)



```



```{r}

#Displaying the PC summary for all three landmark groups
PCA_mean$pc.summary$importance

PCA_basi$pc.summary$importance

PCA_rest$pc.summary$importance


#Does size play a role in the main variation (as it often does)?

rownames(PCA_mean$pc.scores)==names(GPA_Allmean$Csize)
PC_Csize=cbind(PCA_mean$pc.scores[,1:4], log(GPA_Allmean$Csize))
colnames(PC_Csize) <- c("PC1", "PC2", "PC3","PC4","Csize" )

name.check(tree_synch,PC_Csize )
PC_Csize <-as.data.frame(PC_Csize)


PC1_Csize <- gls(PC1~Csize, correlation=corBrownian(1,phy=tree_synch), data=PC_Csize)
PC2_Csize <- gls(PC2~Csize, correlation=corBrownian(1,phy=tree_synch), data=PC_Csize)
PC3_Csize <- gls(PC3~Csize, correlation=corBrownian(1,phy=tree_synch), data=PC_Csize)
PC4_Csize <- gls(PC4~Csize, correlation=corBrownian(1,phy=tree_synch), data=PC_Csize)
PC1_res <-summary(PC1_Csize)
PC2_res<-summary(PC2_Csize)
PC3_res <- summary(PC3_Csize)
PC4_res <- summary(PC4_Csize)


#Allometry graphs
plot.new()
par(mfrow=c(3,1))
par(mar=c(1.2,5,1.2,1.2))

plot(PC1~Csize, data=PC_Csize,pch=19);
abline (lm(PC_Csize$PC1~PC_Csize$Csize));
legend("bottomleft", bty="n",text.font=2,text.col="forest green", cex=1.1,paste("pgls coeff=", round(PC1_res$coefficients[2], 3), "p=",round(PC1_res$tTable[8],3) ))

plot(PC2~Csize, data=PC_Csize, pch=19);
abline (lm(PC_Csize$PC2~PC_Csize$Csize));
legend("bottomleft", bty="n",text.font=2,text.col="forest green", cex=1.1,paste("pgls coeff=", round(PC2_res$coefficients[2], 3), "p=",round(PC2_res$tTable[8],3) ))


plot(PC3~Csize, data=PC_Csize, pch=19);
abline (lm(PC_Csize$PC3~PC_Csize$Csize));
legend("bottomleft",bty="n", text.font=2,text.col="forest green", cex=1.1, paste("pgls coeff=", round(PC3_res$coefficients[2], 3), "p=",round(PC3_res$tTable[8],3) ))


#show_col(viridis_pal()(40)) #Showing the 40 colours present in the viridis package palette



plot(PC1~Csize, data=PC_Csize, type="p",pch=21, bg=col.gp, cex=1.5)
abline (lm(PC_Csize$PC1~PC_Csize$Csize))
text(PC_Csize$PC1~PC_Csize$Csize, labels=Gen.sp, pos=1)
legend(x=10.9,y=-0.125,
legend=c("Caenolestid","Dasyurom.", "Didelphid","Macropod",  expression(italic("Dromiciops")), "Peramelem.","Petaroid", "Phalangeroid",
"Vombatiform"),
pt.bg=c(unique(aggregate(col.gp ~ speclist$Clade, FUN=unique)[,-1]),"light grey"),
pch=c(rep(21,11),24), cex=1, pt.cex=1, ncol = 3, bg= "light grey", box.col="light grey",
text.width = 0.52, y.intersp = 1.3)


#Graphs with just points

par(mfrow=c(3,1))

plot(PC1~Csize, data=PC_Csize, type="p",pch=19, col=colclade)
abline (lm(PC_Csize$PC1~PC_Csize$Csize))


plot(PC2~Csize, data=PC_Csize, type="p",pch=19, col=viridis(1))
abline (lm(PC_Csize$PC2~PC_Csize$Csize))


plot(PC3~Csize, data=PC_Csize, type="p",pch=19, col=viridis(1))
abline (lm(PC_Csize$PC3~PC_Csize$Csize))


```

##The PCA Heatmap 

```{r}

# Find mean configuration - i.e. mean shape in dataset
shape <- GPA_AllSpecimens
consensus <- (select.procrustes(shape, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
diff.from.mean <- coordinates.difference(coordinates = shape$coo, reference = consensus, type = "spherical")

# Ordinate the data
twoD.shape <- two.d.array(shape$coords)
ordination <- stats::prcomp(twoD.shape)

# Force the symmetric component of shape into class "gpagen" (required for variation.range)
gpagen.shape <- list()
gpagen.shape$coords <- shape$coords
gpagen.shape$consensus <- consensus
class(gpagen.shape) <- "gpagen"

# Measure extremes of variation from mean on PC1
PC1.var <- variation.range(gpagen.shape, return.ID = FALSE, axis = 1, ordination = ordination, type = "spherical")

PC2.var <- variation.range(gpagen.shape, return.ID = FALSE, axis = 2, ordination = ordination, type = "spherical")

# Wrap specimens on the tangent space
wrap.PCA <- plotTangentSpace(shape$coords, verbose = FALSE)

# Select extreme specimens (the ones that make the warp-meshes)
hypothetical.1 <- wrap.PCA$pc.shapes[[1]]
hypothetical.2 <- wrap.PCA$pc.shapes[[2]]
hypothetical.3 <- wrap.PCA$pc.shapes[[3]]

```

## Plot the range of variation along PC1 using a heat color scheme
"3d" commands allow precise control of 3D visualization to get dorsal and lateral views

```{r}
setwd("../Data")

# Dorsal view
open3d(windowRect = c(0, 0, 600, 600))
PC1.var.plot <- procrustes.var.plot(hypothetical.1, hypothetical.2, col = viridis, col.val = PC1.var[, "radius"], labels = F)
view3d(phi = 0)  # flip cranium
rgl.snapshot("../PC1heatmap_dor1.png")  # save screenshot for multipanel plot

# Lateral view
open3d(windowRect = c(0, 0, 600, 600))
PC1.var.plot <- procrustes.var.plot(hypothetical.1, hypothetical.2, col = viridis, col.val = PC1.var[, "radius"], labels = F)
view3d(phi = 90)
rgl.snapshot("../PC1heatmap_lat1.png")

#range of variation along PC2
open3d(windowRect = c(0, 0, 600, 600))
PC2.var.plot <- procrustes.var.plot(hypothetical.2, hypothetical.1, col = inferno, col.val = PC2.var[, "radius"], labels = F)
view3d(phi = 0)  # flip cranium
rgl.snapshot("../PC2heatmap_dor.png")  # save screenshot for multipanel plot

#range of variation along PC3
open3d(windowRect = c(0, 0, 600, 600))
PC3.var.plot <- procrustes.var.plot(hypothetical.3, hypothetical.1, col = heat.colors, col.val = PC3.var[, "radius"], labels = F)
view3d(phi = 0)  # flip cranium
rgl.snapshot("../PC2heatmap_dor.png")  # save screenshot for multipanel plot
```

### Prep images for multi-panel figure
This step assumes you have ImageMagik installed on your computer
```{r}
system("mogrify ../Data/Processed/PC1heatmap_*.png -trim ../Data/Processed/*.png")  # removes whitespace (requires ImageMagik)
dor <- readPNG("../Mars-shape/PC1heatmap_dor1.png")
lat <- readPNG("../Mars-shape/PC1heatmap_lat1.png")

```

## Make multipanel Figure 3
ERROR: memory issues when trying to add legend within setEPS. Works when lines 87-104 are run through the console, plot is "zoomed" from RStudio, and then screenshot from that window
```{r}

setEPS()  # sets up plot export
postscript("../Data/Results/Figure3_Heatmaps.eps")
layout(matrix(c(1,1,2,2), 2, 2, byrow = TRUE))  # 2 rows, 2 columns
par(mar=c(0, 0, 0, 0))

# Plot dorsal view
plot(c(0, dim(dor)[2]), c(0, dim(dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) #dim(dor)[2] is the width of the dorsal view, dim(dor)[1] is the height
rasterImage(dor, 0, 0, dim(dor)[2], dim(dor)[1])
text(0, dim(dor)[1]-10, "a", cex = 2)

# Plot lateral view with same dimensions as above
plot(c(0, dim(dor)[2]), c(0, dim(dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T)
rasterImage(lat, 0, 0, dim(dor)[2], dim(dor)[1])
text(0, dim(dor)[1]-10, "b", cex = 2)

# Make gradient legend
legend_image <- as.raster(matrix(viridis(20), ncol = 1))
rasterImage(legend_image, 440, 220, 465, 160)
text(393, 213, "High variation", col = "dark grey", cex = 1.4)
text(420, 168, "Low", col = "dark grey", cex = 1.4)
dev.off()

```


#Warp plots for PC min & max
```{r}
findMeanSpec(GPA_Allmean$coords) # Find specimen closest to mean

MarmurMesh <- read.ply("D:/Downloads/Checkpoint_STL/Marmosa_murina_MVZ197429.ply") # Read in mesh of that specimen

shade3d(MarmurMesh) # Check that mesh is read in

Marmurmshape<-mshape(GPA_Allmean$coords)

plotspec(MarmurMesh,digitspec =  Data[,,25], centered = F) # Check that the landmarks match the mesh

MarmurRefMesh <- warpRefMesh(MarmurMesh, Data[,,25], Marmurmshape, centered=F) # This warps the ref mesh to the mean shape

open3d(); shade3d(MarmurRefMesh); writePLY("D:/Github_Projects/Data/Marmurwarptomean.ply",withColors=T) # Saves mesh to directory



# Warp the refmesh to the min and max of PC1 and PC2 and save mesh to WD

MarmurPC1min <- plotRefToTarget(Marmurmshape, PCA_mean$pc.shapes$PC1min, method="surface",

                              mesh = MarmurRefMesh, mag=1); writePLY("D:/Github_Projects/Data/Results/MarmurPC1Min1.ply",withColors=T)

MarmurPC1max <- plotRefToTarget(Marmurmshape, PCA_mean$pc.shapes$PC1max, method="surface",

                              mesh = MarmurRefMesh, mag=1); writePLY("D:/Github_Projects/Data/Results/MarmurPC1Max1.ply",withColors=T)

MarmurPC2min <- plotRefToTarget(Marmurmshape, PCA_mean$pc.shapes$PC2min, method="surface",

                              mesh = MarmurRefMesh, mag=1); writePLY("D:/Github_Projects/Data/Results/MarmurPC2Min2.ply",withColors=T)

MarmurPC2max <- plotRefToTarget(Marmurmshape, PCA_mean$pc.shapes$PC2max, method="surface",

                              mesh = MarmurRefMesh, mag=1); writePLY("D:/Github_Projects/Data/Results/MarmurPC2Max2.ply",withColors=T)

MarmurPC3min <- plotRefToTarget(Marmurmshape, PCA_mean$pc.shapes$PC3min, method="surface",

                              mesh = MarmurRefMesh, mag=1); writePLY("D:/Github_Projects/Data/Results/MarmurPC3Min3.ply",withColors=T)

MarmurPC31max <- plotRefToTarget(Marmurmshape, PCA_mean$pc.shapes$PC3max, method="surface",

                              mesh = MarmurRefMesh, mag=1); writePLY("D:/Github_Projects/Data/Results/MarmurPC3Max3.ply",withColors=T)

```
