---
title: "Marsupial shape and phylogenetics"
author: "Leonie Lange-Hodgson"
date: "22 January 2019"
output: html_document
editor_options: 
  chunk_output_type: inline
---


# load required libraries

#I will be using the following R packages in these analyses; geomorph, ape, plyr, abind, and geiger.

```{r}
library(rgl)
library(RRPP)
library(geomorph)
library(ape)
library(plyr)
library(abind)
library(geiger)


```
#read in coordinates

#Prior to investigating the three predictions of this study, I will input and prepare my data.
```{r}

#RUN LINE BY LINE! LATER DOWN THE LINE ASK THOMAS HOW TO FIX THIS

#This working directory needs to be set for this chunk because otherwise it won't find the files. It is an issue of rmd talking to R itself.
setwd("E:/Github_Projects/Mars-shape/Data/Raw/Coordinates") #Vera's computer
setwd("D:/Github_Projects/Mars-shape/Data/Raw/Coordinates")  #Leonie's computer


filelist <- list.files(path= "../Data/Raw/Coordinates", pattern = "*.txt")

#Next step is to remove the Museum IDs from the specimen names   
  
  tmp <- matrix(unlist(strsplit(filelist, "_")),ncol=3,byrow = TRUE)   

  filelist_species <- paste(tmp[,1], tmp[,2], sep="_")

  
  names <- gsub (".txt", "", filelist) # extracts names of specimens from the file name
  filelist <- paste("../Data/Raw/Coordinates/", filelist, sep="") # rename with path
  coords <- NULL # make empty object that will be filled with 3D array of coordinate data
  for (i in 1:length(filelist)){
    temp  <- read.morphologika(filelist[i]) 
    k <- dim(temp)[1] 
    coords <- rbind(coords, two.d.array(temp)) }
  Data <- arrayspecs(coords, k, 3) 
  dimnames(Data)[[3]] <- filelist_species
  remove(i, filelist, names, k, coords, temp) # clean up environment
  
  #Double check that the 3D info is entered properly
  plot3d(Data[,,5], asp=FALSE)
  text3d(Data[,,5], texts=c(1:71))
  
 
```

#Continuing data preparation, the function 'gpagen' will run a Generalised Procrustes Analysis (GPA). GPA will take the specimen landmarks and superimpose them, thereby removing all the differences in size, orientation and location between each specimen. What remains are landmark coordinates tha provide the shape variation between species.
#This will make it possible to allow me to analyse their shape variation relative to each other


#GPA for all coordinates
```{r}
#Run GPA
GPA_AllSpecimens <- gpagen(Data)


#AVERAGING MULTIPLE SPECIMENS OF ONE SPECIES

#Two of the species in this dataset have multiple specimens which means they must be averaged prior to analysis. Once this is performed, GPA must be run again to incorporate these changes 

#Remember= dim() is the dimensions of an object

#There are multiple specimens of the same species. This means they need to be averaged before continuing 

# shape data
ind.coords <- aggregate(two.d.array(GPA_AllSpecimens$coords) ~ dimnames(GPA_AllSpecimens$coords)[[3]], FUN=mean)[,-1]
rownames(ind.coords) <- unique(dimnames(GPA_AllSpecimens$coords)[[3]])
ind.coords <- arrayspecs(ind.coords, p=ncol(ind.coords)/3, k=3)
# centroid size data
ind.Csize <- as.vector(aggregate(GPA_AllSpecimens$Csize ~ dimnames(GPA_AllSpecimens$coords)[[3]], FUN=mean)[,-1])
names(ind.Csize) <- unique(names(GPA_AllSpecimens$Csize))
# ind.Csize - These are the individuals Csize data for plotting specimen allometry graphs

#are dimnames identical i.e. will the aligned properly?
dimnames(GPA_AllSpecimens$coords)[[3]]==names(GPA_AllSpecimens$Csize)

dimnames(ind.coords)[[3]]==names(ind.Csize)

#Turning into gdf frame and giving names
GPA_mean <-geomorph.data.frame(coords=ind.coords, Csize=ind.Csize)
dimnames(GPA_mean$coords)[[3]] <-dimnames(ind.coords)[[3]]
names(GPA_mean$Csize) <- names(ind.Csize)


```
#run GPA for basicranial lms
```{r}
#Subset basicranial lms
Data_basi=Data[which(part.gp==2),,]
#double-check that the correct number of landmarks has been subtracted
attributes(Data_basi)

KEEP GOING BELOW TO DO GPA AND AVERAGE MULTIPLE SPECIMENS THEN DO SEPARATE ALLOMETRY PHYSIGNAL, DIET AND LOCOMOTION PGLS ANALYSES.



GPA_AllSpecimens <- gpagen(Data)


#AVERAGING MULTIPLE SPECIMENS OF ONE SPECIES

#Two of the species in this dataset have multiple specimens which means they must be averaged prior to analysis. Once this is performed, GPA must be run again to incorporate these changes 

#Remember= dim() is the dimensions of an object

#There are multiple specimens of the same species. This means they need to be averaged before continuing 

# shape data
ind.coords <- aggregate(two.d.array(GPA_AllSpecimens$coords) ~ dimnames(GPA_AllSpecimens$coords)[[3]], FUN=mean)[,-1]
rownames(ind.coords) <- unique(dimnames(GPA_AllSpecimens$coords)[[3]])
ind.coords <- arrayspecs(ind.coords, p=ncol(ind.coords)/3, k=3)
# centroid size data
ind.Csize <- as.vector(aggregate(GPA_AllSpecimens$Csize ~ dimnames(GPA_AllSpecimens$coords)[[3]], FUN=mean)[,-1])
names(ind.Csize) <- unique(names(GPA_AllSpecimens$Csize))
# ind.Csize - These are the individuals Csize data for plotting specimen allometry graphs

#are dimnames identical i.e. will the aligned properly?
dimnames(GPA_AllSpecimens$coords)[[3]]==names(GPA_AllSpecimens$Csize)

dimnames(ind.coords)[[3]]==names(ind.Csize)

#Turning into gdf frame and giving names
GPA_mean <-geomorph.data.frame(coords=ind.coords, Csize=ind.Csize)
dimnames(GPA_mean$coords)[[3]] <-dimnames(ind.coords)[[3]]
names(GPA_mean$Csize) <- names(ind.Csize)







#Read in classifier file that contains the species name, locomotion type, diet, and phylogenetic clade - 


```{r}

speclist<- read.delim("../Data/Raw/Mars_classifier_list.txt")
rownames(speclist) <-speclist$Specimen

```

#To understand the main variation in shape between marsupial species I will run a Principle Component Analysis (PCA) on the coordinate data. Using the function plotTangentSpace from the geomorph package, the resulting graph will provide the two PC axes that explain the most variation between species coordinates. To understand the variance of each PC axis, I will run the function pc.sumary$importance to view the eigenvalues. These eigenvalues aid in further analyses.

#Each PC represents a particular aspect of the overall shape variation, see more in the chunk

##would you do a PCA for habitat, diet and clade? 

```{r}
#Checking if all specimens are read in properly with their Csizes
PCA_all <- plotTangentSpace(GPA_AllSpecimens$coords, label=dimnames(GPA_AllSpecimens$coords)[[3]])

plot(PCA_all$pc.scores[,1]~GPA_AllSpecimens$Csize ); text(PCA_all$pc.scores[,1]~GPA_AllSpecimens$Csize, labels=names(GPA_AllSpecimens$Csize))


#Run PCA
PCA_mean <- plotTangentSpace(GPA_mean$coords, label=dimnames(GPA_mean$coords)[[3]])

plot(PCA_mean$pc.scores[,1]~GPA_mean$Csize ); text(PCA_mean$pc.scores[,1]~GPA_mean$Csize, labels=names(GPA_mean$Csize))

#If there is an obvious problem specimen, find out what number specimen it is on the GPA
dimnames(GPA_mean$coords)[[3]]

plotRefToTarget(GPA_mean$coords[,,1], GPA_mean$coords[,,48], method="vector")


```

#PlotRefto target - Function plots shape differences between a reference and target specimen
#vector: a plot showing the vector displacements between corresponding landmarks in the reference and target specimen is shown


# Visualize main differences OF WHAT AND WHY? YOU MIGHT ALSO ADD SOME DESCRIPTION OF THE RESULTS HERE, WHICH YOU CAN LATER PUT INTO RESULTS SECTION.
```{r}

#Below is a starting point for customizing your PlotRefToTarget plots. I recommend playing with this! This then gets fed into plotRefToTarget as per below.
PRTT=gridPar(pt.bg="green", pt.size = 1)

plotRefToTarget(PCA_mean$pc.shapes$PC1min, PCA_mean$pc.shapes$PC1max, method="vector", mag=0.6,gridPars = PRTT)
plotRefToTarget(PCA_mean$pc.shapes$PC2min, PCA_mean$pc.shapes$PC2max, method="vector") #Back to standard grey
```

#TREE PHYLOGENY 

#names(tree_use$tip.label[NUMBER] <- "New_Species") #What I need to write to change the name of a particular tip

```{r}
#All 3 analyses (Maximum likelihood (ML), Bayesian, and parsimony) performed in Mitchell et al. (2014) resulted in well-resolved and concordant trees.
#I chose the first tree (ML, without outgroups) due to the use of it in their further analyses

tree=read.nexus("../Data/Raw/Mitchell_pruned_tree.nex")
tree_use <- tree[1]$tree_1

###~~~~~~~CHANGING SPECIES NAME~~~~~~~~~~###
tree_use$tip.label
tree_use$tip.label[5]  
#This shows particular tip label (without the number present this code will provide you with all the tip labels)

tree_use$tip.label[5] <- "Caluromys_derbianus"
tree_use$tip.label[1] <- "Lestoros_inca"
tree_use$tip.label[178] <- "Dendrolagus_inustus"
tree_use$tip.label[30] <- "Philander_andersoni"
tree_use$tip.label[124] <- "Phalanger_mimicus"
tree_use$tip.label[66] <- "Planigale_ingrami"
tree_use$tip.label[67] <- "Sminthopsis_murina"
tree_use$tip.label[39] <- "Thylamys_elegans"
tree_use$tip.label[20] <- "Marmosa_demerarae"
tree_use$tip.label[19] <- "Marmosa_alstoni"
tree_use$tip.label[130] <- "Petaurus_australis"
#to do the namecheck, Csize is easier to navigate to but it is part of the same gdf so that's fine

#NAVIGATE TO THE NAMES FOR THESE NAME CHECKS HAVE A LOOK FURTHER UP IN THE CODE
namecheck=name.check(tree_use,GPA_mean$Csize)

Matchtest=match(tree_use$tip.label,GPA_mean$Csize)
length(which(!is.na (Matchtest)))


tree_synch=drop.tip(tree_use,namecheck$tree_not_data)

name.check(tree_synch,GPA_mean$Csize)

```

#Create two partitions

The landmark dataset will now be split into two partitions; basicranium and the rest of the skull. 


#IF YOU RUN INTO A PROBLEM, FIRST CHECK IF ALL THE DATA HAVE READ IN PROPERLY - THERE WERE TWO ISSUES WITH PART.GP, ONE WAS THAT THE FIRST ROW CONTAINED WEIRD CHARACTERS AND THE SECOND WAS THAT IT CONTAINED HEAPS OF COLUMNS WITH JUST NAS. BOTH ARE FIXED NOW
```{r}
### Import the  partition map, ensure it's a factor
part.gp=as.vector(read.csv("../Data/Raw/partitions.csv", header=FALSE))
part.gp=as.factor(part.gp$V1)
#Creating the two partitions 
  
basi <- GPA_mean$coords[which(part.gp==2), , ]
basiCsize <- GPA_mean$Csize

basiland <-geomorph.data.frame(coords=basi, Csize=basiCsize)
dimnames(basiland$coords)[[3]] <-dimnames(GPA_mean$coords)[[3]]
names(basiland$Csize) <- names(GPA_mean$Csize)

#Must return all TRUE
names(basiland$Csize)==dimnames(basiland$coords)[[3]]

reskull <-GPA_mean$coords[which(part.gp==1), , ]



```
Temporary rough code to make partitions show up 


```{r}
#colouring partition spheres CHANGE THESE TO WHAT IT IS HERE SO YOU KNOW WHAT IS GOING ON

 #provides the numbers of the parts
  PartLevels = unique(part.gp[])
  #colour scheme ;-)
  Colours <- c("hotpink", "purple")
 #Make mean shape that the points can be plotted on
  Ref=mshape(GPA_mean$coords)
  
  Part=list()
  ##subset the landmarks according to the partitions
  for(i in 1:length(PartLevels)){
    Part[[i]]<-which (part.gp[] == PartLevels[[i]])
  }
  
  ##colours the spheres for each partition on the mean shape
  
  open3d()
  
  for (i in 1:length(PartLevels)){
    spheres3d( Ref[Part[[i]],], col=Colours[i], lit=TRUE, radius = 0.003, asp=F)
    
  }

  
```

Allometry in geometric morphometrics refers to size-related changes of morphological traits. 
e.g. does brain size increase as body size increases?
looking at overall shape however

#TO UNDERSTAND/TEST...., TESTING FOR ALLOMETRY (WHAT ABOUT ALLOMETRY ARE YOU TESTING?)

#Figure out allometry - try to do this yourself, check the geomorph guide here. Is there significant allometry? Can you plot it? Is there significant correlation betwen PC1, 2 or 3 and Allometry? As a proxy for size, use GPA$Csize.Try this, I can help you later tonight if you get stuck.


#CHECK IF ALLOMETRY IS ONLY IN THE "REST OF SKULL" PARTITION

```{r}
#Geomorph update 3.1.0 on the 27th March removed procD.Allometry so using the function procD.lm instead. 
Allom<-procD.lm(coords~Csize, logsz=TRUE, data=GPA_mean, iter=10000, print.progress=FALSE)
summary(Allom)

summary(procD.lm(coords~Csize,  data=GPA_mean, logsz=TRUE,iter=10000, print.progress=TRUE))


AllBasi<-procD.lm(coords~Csize, data=basiland, iter=1000, print.progress=FALSE)
summary(AllBasi)

```



#Compare evolutionary rates of partitions and phylogenetic signal

There are four partitions but you only set up 3 of these res.evolrate codes (see your code). Do I use 1 if I have 2 partitions?
```{r}
#### SHAPE RATE ANALYSES
### Compare evolutionary rates between partitions
# Calculate rates for each partition


parevo.rate <- 
  compare.multi.evol.rates(GPA_mean$coords,  part.gp, tree_synch,iter = 1000)

# Does phylogenetic signal differ? do for the other two as well and csize

physignal(GPA_mean$coords, tree_synch, iter = 1000)



```

#
