---
title: "Marsupial shape and phylogenetics"
author: "Leonie Lange-Hodgson"
date: "22 January 2019"
output: html_document
editor_options: 
  chunk_output_type: inline
---


# load required libraries

#I will be using the following R packages in these analyses; geomorph, ape, plyr, abind, and geiger.

```{r}

# Note: need to have dispRity ticked in the library before you can install landVR

library(rgl)
library(RRPP)
library(geomorph)
library(ape)
library(plyr)
library(abind)
library(geiger)
library(caper)
library(devtools)
library(phytools)
library(nlme)
library(dispRity)
library(landvR)
library(png)

Dviv <- read.ply("../Data/Dasyurus_viverrinus_ascii.ply")


```
#read in coordinates

#Prior to investigating the three predictions of this study, I will input and prepare my data.
```{r}

#RUN LINE BY LINE! LATER DOWN THE LINE ASK THOMAS HOW TO FIX THIS

#This working directory needs to be set for this chunk because otherwise it won't find the files. It is an issue of rmd talking to R itself.
setwd("E:/Github_Projects/Mars-shape/Data/Raw/Coordinates") #Vera's computer
setwd("D:/Github_Projects/Mars-shape/Data/Raw/Coordinates")  #Leonie's computer


filelist <- list.files(path= "../Data/Raw/Coordinates", pattern = "*.txt")

#Next step is to remove the Museum IDs from the specimen names   
  
  tmp <- matrix(unlist(strsplit(filelist, "_")),ncol=3,byrow = TRUE)   

  filelist_species <- paste(tmp[,1], tmp[,2], sep="_")

  
  names <- gsub (".txt", "", filelist) # extracts names of specimens from the file name
  filelist <- paste("../Data/Raw/Coordinates/", filelist, sep="") # rename with path
  coords <- NULL # make empty object that will be filled with 3D array of coordinate data
  for (i in 1:length(filelist)){
    temp  <- read.morphologika(filelist[i]) 
    k <- dim(temp)[1] 
    coords <- rbind(coords, two.d.array(temp)) }
  Data <- arrayspecs(coords, k, 3) 
  dimnames(Data)[[3]] <- filelist_species
  remove(i, filelist, names, k, coords, temp) # clean up environment
  
  #Double check that the 3D info is entered properly
  plot3d(Data[,,4], asp=FALSE)
  text3d(Data[,,4], texts=c(1:58))
  
 
```

#Read in classifier file that contains the species name, locomotion type, diet, and clade


```{r}

speclist<- read.delim("../Data/Raw/Mars_classifier_list.txt")
rownames(speclist) <-speclist$Specimen

```

#Importation of partition map, ensure it's a factor


The landmark dataset will now be split into two partitions; basicranium and the rest of the skull. 

```{r}

### Import the  partition map, ensure it's a factor
part.gp=as.vector(read.csv("../Data/Raw/partitions.csv", header=FALSE))
part.gp=as.factor(part.gp$V1)


```

# PHYLOGENY 

#names(tree_use$tip.label[NUMBER] <- "New_Species") #What I need to write to change the name of a particular tip

```{r}
#All 3 analyses (Maximum likelihood (ML), Bayesian, and parsimony) performed in Mitchell et al. (2014) resulted in well-resolved and concordant trees.
#I chose the first tree (ML, without outgroups) due to the use of it in their further analyses

tree=read.nexus("../Data/Raw/Mitchell_pruned_tree.nex")
tree_use <- tree[1]$tree_1

###~~~~~~~CHANGING SPECIES NAME~~~~~~~~~~###
tree_use$tip.label
tree_use$tip.label[5]  
#This shows particular tip label (without the number present this code will provide you with all the tip labels)

tree_use$tip.label[5] <- "Caluromys_derbianus"
tree_use$tip.label[1] <- "Lestoros_inca"
tree_use$tip.label[178] <- "Dendrolagus_inustus"
tree_use$tip.label[30] <- "Philander_andersoni"
tree_use$tip.label[124] <- "Phalanger_mimicus"
tree_use$tip.label[66] <- "Planigale_ingrami"
tree_use$tip.label[67] <- "Sminthopsis_murina"
tree_use$tip.label[39] <- "Thylamys_elegans"
tree_use$tip.label[20] <- "Marmosa_demerarae"
tree_use$tip.label[19] <- "Marmosa_alstoni"
tree_use$tip.label[130] <- "Petaurus_australis"



```

#Continuing data preparation, the function 'gpagen' will run a Generalised Procrustes Analysis (GPA). GPA will take the specimen landmarks and superimpose them, thereby removing all the differences in size, orientation and location between each specimen. What remains are landmark coordinates tha provide the shape variation between species.
#This will make it possible to allow me to analyse their shape variation relative to each other


#GPA for all coordinates
```{r}
#Run GPA
GPA_AllSpecimens <- gpagen(Data)


#AVERAGING MULTIPLE SPECIMENS OF ONE SPECIES

#Two of the species in this dataset have multiple specimens which means they must be averaged prior to analysis. Once this is performed, GPA must be run again to incorporate these changes 

#Remember= dim() is the dimensions of an object


# shape data
ind.coords <- aggregate(two.d.array(GPA_AllSpecimens$coords) ~ dimnames(GPA_AllSpecimens$coords)[[3]], FUN=mean)[,-1]
rownames(ind.coords) <- unique(dimnames(GPA_AllSpecimens$coords)[[3]])
ind.coords <- arrayspecs(ind.coords, p=ncol(ind.coords)/3, k=3)
# centroid size data
ind.Csize <- as.vector(aggregate(GPA_AllSpecimens$Csize ~ dimnames(GPA_AllSpecimens$coords)[[3]], FUN=mean)[,-1])
names(ind.Csize) <- unique(names(GPA_AllSpecimens$Csize))
# ind.Csize - These are the individuals Csize data for plotting specimen allometry graphs

#are dimnames identical i.e. will the aligned properly?
dimnames(GPA_AllSpecimens$coords)[[3]]==names(GPA_AllSpecimens$Csize)

dimnames(ind.coords)[[3]]==names(ind.Csize)

#Turning into gdf frame and giving names
GPA_Allmean <-geomorph.data.frame(coords=ind.coords, Csize=ind.Csize)
dimnames(GPA_Allmean$coords)[[3]] <-dimnames(ind.coords)[[3]]
names(GPA_Allmean$Csize) <- names(ind.Csize)

#cHECK that it works with phylogeny
namecheck=name.check(tree_use,GPA_Allmean$Csize)

#Checking to see if the species coordinates and the species on the tree match
Matchtest=match(tree_use$tip.label,GPA_Allmean$Csize)
length(which(!is.na (Matchtest)))

#Removing all species in the tree that is not apart of the study
tree_synch=drop.tip(tree_use,namecheck$tree_not_data)

name.check(tree_synch,GPA_Allmean$Csize)


```


#GPA for basicranial landmarks
```{r}
#Subset basicranial landmarks
Data_basi=Data[which(part.gp==2),,]
#double-check that the correct number of landmarks has been subtracted
attributes(Data_basi)

#GPA
GPA_basi <- gpagen(Data_basi)


#AVERAGING MULTIPLE SPECIMENS OF ONE SPECIES

#Two of the species in this dataset have multiple specimens which means they must be averaged prior to analysis. Once this is performed, GPA must be run again to incorporate these changes 

#Remember= dim() is the dimensions of an object


# shape data
basi.coords <- aggregate(two.d.array(GPA_basi$coords) ~ dimnames(GPA_basi$coords)[[3]], FUN=mean)[,-1]
rownames(basi.coords) <- unique(dimnames(GPA_basi$coords)[[3]])
basi.coords <- arrayspecs(basi.coords, p=ncol(basi.coords)/3, k=3)
# centroid size data
basi.Csize <- as.vector(aggregate(GPA_basi$Csize ~ dimnames(GPA_basi$coords)[[3]], FUN=mean)[,-1])
names(basi.Csize) <- unique(names(GPA_basi$Csize))
# basi.Csize - These are the individuals Csize data for plotting specimen allometry graphs

#are dimnames identical i.e. will the aligned properly?
dimnames(GPA_basi$coords)[[3]]==names(GPA_basi$Csize)

dimnames(basi.coords)[[3]]==names(basi.Csize)

#Turning into gdf frame and giving names
GPA_Basimean <-geomorph.data.frame(coords=basi.coords, Csize=basi.Csize)
dimnames(GPA_Basimean$coords)[[3]] <-dimnames(basi.coords)[[3]]
names(GPA_Basimean$Csize) <- names(basi.Csize)


```



#GPA for the REST OF THE SKULL


```{r}

#Subset the rest of the skull's landmarks lms
Data_reskull=Data[which(part.gp==1),,]
#double-check that the correct number of landmarks has been subtracted
attributes(Data_reskull)

#GPA

GPA_reskull <- gpagen(Data_reskull)


#AVERAGING MULTIPLE SPECIMENS OF ONE SPECIES

#Two of the species in this dataset have multiple specimens which means they must be averaged prior to analysis. Once this is performed, GPA must be run again to incorporate these changes 

#Remember= dim() is the dimensions of an object

# shape data
ind.coords <- aggregate(two.d.array(GPA_reskull$coords) ~ dimnames(GPA_reskull$coords)[[3]], FUN=mean)[,-1]
rownames(ind.coords) <- unique(dimnames(GPA_reskull$coords)[[3]])
ind.coords <- arrayspecs(ind.coords, p=ncol(ind.coords)/3, k=3)
# centroid size data
ind.Csize <- as.vector(aggregate(GPA_reskull$Csize ~ dimnames(GPA_reskull$coords)[[3]], FUN=mean)[,-1])
names(ind.Csize) <- unique(names(GPA_reskull$Csize))
# ind.Csize - These are the individuals Csize data for plotting specimen allometry graphs

#are dimnames identical i.e. will the aligned properly?
dimnames(GPA_reskull$coords)[[3]]==names(GPA_reskull$Csize)

dimnames(ind.coords)[[3]]==names(ind.Csize)

#Turning into gdf frame and giving names
GPA_Restmean <-geomorph.data.frame(coords=ind.coords, Csize=ind.Csize)
dimnames(GPA_Restmean$coords)[[3]] <-dimnames(ind.coords)[[3]]
names(GPA_Restmean$Csize) <- names(ind.Csize)



```


#Images for partitions


```{r}
#colouring partition spheres 

 #provides the numbers of the parts
  PartLevels = unique(part.gp[])
  #colour scheme ;-)
  Colours <- c("hotpink", "purple")
  
  #Mesh material
  
 
  #Make mean shape that the points can be plotted on
  Ref=mshape(GPA_Allmean$coords)
  
  #Or, for 3d plotting, call up the coordinates of the mean specimen; remember that the radius of the spheres might need to increase because this is not GPA'ed
  Ref <- Data[,,7]
  
  Part=list()
  ##subset the landmarks according to the partitions
  for(i in 1:length(PartLevels)){
    Part[[i]]<-which (part.gp[] == PartLevels[[i]])
  }
  
  ##colours the spheres for each partition on the mean shape
  
  open3d()
  
  for (i in 1:length(PartLevels)){
    spheres3d( Ref[(Part[[i]]),]*1000, col=Colours[i], lit=TRUE, radius = 900, asp=F)
    
  }

  shade3d (Dviv, color="grey")
  
```


#Evolutionary allometry

```{r}
#Geomorph update 3.1.0 on the 27th March removed procD.Allometry so using the function procD.lm instead. 
Phyall<-procD.pgls(coords~Csize, tree_synch, data=GPA_Allmean, iter=999)
summary(Phyall)


Phyallbasi<-procD.pgls(coords~Csize, tree_synch, data=GPA_Basimean, iter=999)
summary(Phyallbasi)

plotAllometry(Phyallbasi, size=GPA_Basimean$Csize, method = "RegScore")

Phyallrest<-procD.pgls(coords~Csize, tree_synch, data=GPA_Restmean, iter=999)
summary(Phyallrest)


```


#PlotRefto target - Function plots shape differences between a reference and target specimen
#vector: a plot showing the vector displacements between corresponding landmarks in the reference and target specimen is shown


# Visualize main differences OF WHAT AND WHY? YOU MIGHT ALSO ADD SOME DESCRIPTION OF THE RESULTS HERE, WHICH YOU CAN LATER PUT INTO RESULTS SECTION.
```{r}

#Below is a starting point for customizing your PlotRefToTarget plots. I recommend playing with this! This then gets fed into plotRefToTarget as per below.
PRTT=gridPar(pt.bg="green", pt.size = 1)

plotRefToTarget(PCA_mean$pc.shapes$PC1min, PCA_mean$pc.shapes$PC1max, method="vector", mag=0.6,gridPars = PRTT)
plotRefToTarget(PCA_mean$pc.shapes$PC2min, PCA_mean$pc.shapes$PC2max, method="vector") #Back to standard grey
```



Allometry in geometric morphometrics refers to size-related changes of morphological traits. 
e.g. does brain size increase as body size increases?
looking at overall shape however

#Csize has an effect on shape of the landmarks

##When analysing the allometry of the remaining landmark on the skull the results showed the effect size of Csize is quite small however there is a fairly significant p-value suggesting that Csize does have an effect on shape

```{r}
#Geomorph update 3.1.0 on the 27th March removed procD.Allometry so using the function procD.lm instead. 
Allom<-procD.lm(coords~Csize, logsz=TRUE, data=GPA_Allmean, iter=10000, print.progress=FALSE)
summary(Allom)


AllBasi<-procD.lm(coords~Csize, data=GPA_Basimean, iter=1000, print.progress=FALSE)
summary(AllBasi)

plotAllometry(AllBasi, size=GPA_Basimean$Csize, method = "RegScore")

AllRest<-procD.lm(coords~Csize, logsz=TRUE, data=GPA_Restmean, iter=10000, print.progress=FALSE)
summary(AllRest)

```



#Compare evolutionary rates of partitions and phylogenetic signal



Evolutionary rates

#Phylogenetic signalling


```{r}
#### SHAPE RATE ANALYSES
### Compare evolutionary rates between partitions
# Calculate rates for each partition


parevo.rate <- 
  compare.multi.evol.rates(GPA_Allmean$coords,  part.gp, tree_synch,iter = 1000)
summary(parevo.rate)

# Does phylogenetic signal differ? do for the other two as well and csize

physignal(GPA_Allmean$coords, tree_synch, iter = 1000)

physignal(GPA_Basimean$coords, tree_synch, iter = 1000)

physignal(GPA_Restmean$coords, tree_synch, iter = 1000)

```

#PGLS for Diet and Locomotion (look at all three GPAs)


```{r}
#Combine the GPA geomorph dataframe with the two columns I want from the speclist data frame

#Do this for each of the GPA dataframes


Allandlist<-mutate(GPA_Allmean, diet=speclist$Diet, locomotion=speclist$Locomotion)

Basiandlist<-mutate(GPA_Basimean, diet=speclist$Diet, locomotion=speclist$Locomotion)

Restandlist<-mutate(GPA_Restmean, diet=speclist$Diet, locomotion=speclist$Locomotion)


#First look at diet and locomotion for all landmarks


Alldiet_inter<-procD.pgls(coords ~ diet, tree_synch, data = Allandlist, iter = 999)
summary(Alldiet_inter)

AllLoco_add<-procD.pgls(coords ~ locomotion, tree_synch, data = Allandlist, iter=999)
summary(AllLoco_add)


# Next look at basicranium landmarks

Basidiet_inter<-procD.pgls(coords~diet, tree_synch, data = Basiandlist, iter = 999)
summary(Basidiet_inter)

BasiLoco_inter<-procD.pgls(coords~locomotion, tree_synch, data = Basiandlist, iter=999)
summary(BasiLoco_inter)

plot3d(Basiandlist$coords[,,8], asp=TRUE); text3d(Basiandlist$coords[,,8], texts=c(35:58))

                                            
                                                   
#Then at the rest of the skull landmarks

RestDiet_inter<-procD.pgls(coords~diet, tree_synch, data = Restandlist, iter = 999)
summary(RestDiet_inter)

RestLoco_add<-procD.pgls(coords~locomotion, tree_synch, data = Restandlist, iter=999)
summary(RestLoco_add)


```


#To understand the main variation in shape between marsupial species I run a Principal Component Analysis (PCA) on the coordinate data. Using the function plotTangentSpace from the geomorph package, the resulting graph will provide the two PC axes that explain the most variation between species coordinates. To understand the variance of each PC axis, I will run the function pc.sumary$importance to view the eigenvalues. These eigenvalues aid in further analyses.

#Each PC represents a particular aspect of the overall shape variation, see more in the chunk

```{r}
#Entering this code to allow me to see the positive and negative extreme ends of PC2
PCA_m23 <- plotTangentSpace(GPA_Allmean$coords, axis1 = 2, axis2 = 3, label=dimnames(GPA_Allmean$coords)[[3]])

PCA_m31 <- plotTangentSpace(GPA_Allmean$coords, axis1 = 3, axis2 = 1, label=dimnames(GPA_Allmean$coords)[[3]])

#Run PCA for all three partitions
PCA_mean <- plotTangentSpace(GPA_Allmean$coords, label=dimnames(GPA_Allmean$coords)[[3]])

PCA_basi <- plotTangentSpace(GPA_Basimean$coords, label=dimnames(GPA_Basimean$coords)[[3]])

PCA_rest <- plotTangentSpace(GPA_Restmean$coords, label=dimnames(GPA_Restmean$coords)[[3]])

#Running another PCA that displays the results in morphospace
plotGMPhyloMorphoSpace(tree_synch, GPA_Allmean$coords )

#Displaying the PC summary
PCA_mean$pc.summary$importance


#Does size play a role in the main variation (as it often does)?

rownames(PCA_mean$pc.scores)==names(GPA_Allmean$Csize)
PC_Csize=cbind(PCA_mean$pc.scores[,1:4], log(GPA_Allmean$Csize))
colnames(PC_Csize) <- c("PC1", "PC2", "PC3","PC4","Csize" )

name.check(tree_synch,PC_Csize )
PC_Csize <-as.data.frame(PC_Csize)


PC1_Csize <- gls(PC1~Csize, correlation=corBrownian(1,phy=tree_synch), data=PC_Csize)
PC2_Csize <- gls(PC2~Csize, correlation=corBrownian(1,phy=tree_synch), data=PC_Csize)
PC3_Csize <- gls(PC3~Csize, correlation=corBrownian(1,phy=tree_synch), data=PC_Csize)
PC4_Csize <- gls(PC4~Csize, correlation=corBrownian(1,phy=tree_synch), data=PC_Csize)
summary(PC1_Csize)
summary(PC2_Csize)
summary(PC3_Csize)
summary(PC4_Csize)

par(mfrow=c(2,2))

plot(PC1~Csize, data=PC_Csize)
abline (lm(PC_Csize$PC1~PC_Csize$Csize))
text(PC_Csize$PC1~PC_Csize$Csize, labels=rownames(PC_Csize))

plot(PC2~Csize, data=PC_Csize)
abline (lm(PC_Csize$PC2~PC_Csize$Csize))
text(PC_Csize$PC2~PC_Csize$Csize, labels=rownames(PC_Csize))

plot(PC3~Csize, data=PC_Csize)
abline (lm(PC_Csize$PC3~PC_Csize$Csize))
text(PC_Csize$PC3~PC_Csize$Csize, labels=rownames(PC_Csize))

plot(PC4~Csize, data=PC_Csize)
abline (lm(PC_Csize$PC4~PC_Csize$Csize))
text(PC_Csize$PC4~PC_Csize$Csize, labels=rownames(PC_Csize))


#Checking phylogenetic signal in PC1; add justification of K reference in thesis; think about if we want this in because it is very different to shape physig.

phylosig(tree_synch, PCA_mean$pc.scores[,1],method="K", nsim=1000)
phylosig(tree_synch, PCA_mean$pc.scores[,2],method="K", nsim=1000)
phylosig(tree_synch, PCA_mean$pc.scores[,3],method="K", nsim=1000)
phylosig(tree_synch, PCA_mean$pc.scores[,4],method="K", nsim=1000)

```

##The PCA Heatmap 

```{r}

# Find mean configuration - i.e. mean shape in dataset
shape <- GPA_AllSpecimens
consensus <- (select.procrustes(shape, selector = mean))[[1]]

# Measure spherical coordinates differences from the mean for each specimen for each landmark
diff.from.mean <- coordinates.difference(coordinates = shape$coo, reference = consensus, type = "spherical")

# Ordinate the data
twoD.shape <- two.d.array(shape$coords)
ordination <- stats::prcomp(twoD.shape)

# Force the symmetric component of shape into class "gpagen" (required for variation.range)
gpagen.shape <- list()
gpagen.shape$coords <- shape$coords
gpagen.shape$consensus <- consensus
class(gpagen.shape) <- "gpagen"

# Measure extremes of variation from mean on PC1
PC1.var <- variation.range(gpagen.shape, return.ID = FALSE, axis = 1, ordination = ordination, type = "spherical")

# Wrap specimens on the tangent space
wrap.PCA <- plotTangentSpace(shape$coords, verbose = FALSE)

# Select extreme specimens (the ones that make the warp-meshes)
hypothetical.1 <- wrap.PCA$pc.shapes[[1]]
hypothetical.2 <- wrap.PCA$pc.shapes[[2]]


```

## Plot the range of variation along PC1 using a heat color scheme
"3d" commands allow precise control of 3D visualization to get dorsal and lateral views

```{r}
setwd("../Data")

# Dorsal view
open3d(windowRect = c(0, 0, 600, 600))
PC1.var.plot <- procrustes.var.plot(hypothetical.1, hypothetical.2, col = heat.colors, col.val = PC1.var[, "radius"], labels = F)
view3d(phi = 0)  # flip cranium
rgl.snapshot("../PC1heatmap_dor.png")  # save screenshot for multipanel plot

# Lateral view
open3d(windowRect = c(0, 0, 600, 600))
PC1.var.plot <- procrustes.var.plot(hypothetical.1, hypothetical.2, col = heat.colors, col.val = PC1.var[, "radius"], labels = F)
view3d(phi = 90)
rgl.snapshot("../PC1heatmap_lat.png")
```

### Prep images for multi-panel figure
This step assumes you have ImageMagik installed on your computer
```{r}
system("mogrify ../Data/Processed/PC1heatmap_*.png -trim ../Data/Processed/*.png")  # removes whitespace (requires ImageMagik)
dor <- readPNG("../Data/Processed/PC1heatmap_dor.png")
lat <- readPNG("../Data/Processed/PC1heatmap_lat.png")

```

## Make multipanel Figure 3
ERROR: memory issues when trying to add legend within setEPS. Works when lines 87-104 are run through the console, plot is "zoomed" from RStudio, and then screenshot from that window
```{r}

setEPS()  # sets up plot export
postscript("../Data/Results/Figure3_Heatmaps.eps")
layout(matrix(c(1,1,2,2), 2, 2, byrow = TRUE))  # 2 rows, 2 columns
par(mar=c(0, 0, 0, 0))

# Plot dorsal view
plot(c(0, dim(dor)[2]), c(0, dim(dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T) #dim(dor)[2] is the width of the dorsal view, dim(dor)[1] is the height
rasterImage(dor, 0, 0, dim(dor)[2], dim(dor)[1])
text(0, dim(dor)[1]-10, "a", cex = 2)

# Plot lateral view with same dimensions as above
plot(c(0, dim(dor)[2]), c(0, dim(dor)[1]), type = "n", axes = F, xlab = "", ylab = "", asp = T)
rasterImage(lat, 0, 0, dim(dor)[2], dim(dor)[1])
text(0, dim(dor)[1]-10, "b", cex = 2)

# Make gradient legend
legend_image <- as.raster(matrix(heat.colors(20), ncol = 1))
rasterImage(legend_image, 440, 220, 465, 160)
text(393, 213, "High variation", col = "dark grey", cex = 1.4)
text(420, 168, "Low", col = "dark grey", cex = 1.4)
dev.off()

```

```{r}
findMeanSpec(GPA_Allmean$coords) # Find specimen closest to mean

MarmurMesh <- read.ply("D:/Github_Projects/Data/Marmosa_murina_MVZ197429.ply") # Read in mesh of that specimen

shade3d(MarmurMesh) # Check that mesh is read in

Marmurmshape<-mshape(GPA_Allmean$coords)

plotspec(MarmurMesh,Data[,,25], centered = F) # Check that the landmarks match the mesh

MarmurRefMesh <- warpRefMesh(MarmurMesh, Data[,,25], Marmurmshape, centered=F) # This warps the ref mesh to the mean shape

open3d(); shade3d(MarmurRefMesh); writePLY("D:/Github_Projects/Data/Marmurwarptomean.ply",withColors=T) # Saves mesh to directory



# Warp the refmesh to the min and max of PC1 and PC2 and save mesh to WD

MarmurPC1min <- plotRefToTarget(Marmurmshape, PCA_mean$pc.shapes$PC1min, method="surface",

                              mesh = MarmurRefMesh, mag=1); writePLY("D:/Github_Projects/Data/MarmurPC1Min1.ply",withColors=T)

MarmurPC1max <- plotRefToTarget(Marmurmshape, PCA_mean$pc.shapes$PC1max, method="surface",

                              mesh = MarmurRefMesh, mag=1); writePLY("D:/Github_Projects/Data/MarmurPC1Max1.ply",withColors=T)

MarmurPC2min <- plotRefToTarget(Marmurmshape, PCA_mean$pc.shapes$PC2min, method="surface",

                              mesh = MarmurRefMesh, mag=1); writePLY("D:/Github_Projects/Data/MarmurPC2Min2.ply",withColors=T)

MarmurPC2max <- plotRefToTarget(Marmurmshape, PCA_mean$pc.shapes$PC2max, method="surface",

                              mesh = MarmurRefMesh, mag=1); writePLY("D:/Github_Projects/Data/MarmurPC2Max2.ply",withColors=T)

MarmurPC3min <- plotRefToTarget(Marmurmshape, PCA_mean$pc.shapes$PC3min, method="surface",

                              mesh = MarmurRefMesh, mag=1); writePLY("D:/Github_Projects/Data/MarmurPC3Min3.ply",withColors=T)

MarmurPC31max <- plotRefToTarget(Marmurmshape, PCA_mean$pc.shapes$PC3max, method="surface",

                              mesh = MarmurRefMesh, mag=1); writePLY("D:/Github_Projects/Data/MarmurPC3Max3.ply",withColors=T)

```
