---
title: "Marsupial shape and phylogenetics"
author: "Leonie Lange-Hodgson"
date: "22 January 2019"
output: html_document
---


# load required libraries

```{r}

library(geomorph)
library(ape)
library(plyr)
library(abind)
library(geiger)

```
#read in coordinates
```{r}

#RUN LINE BY LINE! LATER DOWN THE LINE ASK THOMAS HOW TO FIX THIS

#This working directory needs to be set for this chunk because otherwise it won't find the files. It is an issue of rmd talking to R itself.
setwd("E:/Github_Projects/Mars-shape/Data/Raw/Coordinates") #Vera's computer
setwd("D:/Github_Projects/Mars-shape/Data/Raw/Coordinates")  #Leonie's computer


filelist <- list.files(path= "../Data/Raw/Coordinates", pattern = "*.txt")

#Next step is to remove the Museum IDs from the specimen names   
  
  tmp <- matrix(unlist(strsplit(filelist, "_")),ncol=3,byrow = TRUE)   

  filelist_species <- paste(tmp[,1], tmp[,2], sep="_")

  
  names <- gsub (".txt", "", filelist) # extracts names of specimens from the file name
  filelist <- paste("../Data/Raw/Coordinates/", filelist, sep="") # rename with path
  coords <- NULL # make empty object that will be filled with 3D array of coordinate data
  for (i in 1:length(filelist)){
    temp  <- read.morphologika(filelist[i]) 
    k <- dim(temp)[1] 
    coords <- rbind(coords, two.d.array(temp)) }
  Data <- arrayspecs(coords, k, 3) 
  dimnames(Data)[[3]] <- filelist_species
  remove(i, filelist, names, k, coords, temp) # clean up environment
  
  #Double check that the 3D info is entered properly
  plot3d(Data[,,5], asp=FALSE)
  text3d(Data[,,5], texts=c(1:71))

  
  
```



#GPA - superimposing a group of shapes (removing all the differences in size/scale, orientation and location between this group of species) so that the only differences in the coordinates of corresponding landmarks between species must be the result of differences in shape between those species
```{r}
#Run GPA
GPA <- gpagen(Data)


```



#AVERAGING MULTIPLE SPECIMENS OF ONE SPECIES

```{r}
#There are multiple specimens of the same species. This means they need to be averaged before continuing 

# shape data
ind.coords <- aggregate(two.d.array(GPA$coords) ~ dimnames(GPA$coords)[[3]], FUN=mean)[,-1]
rownames(ind.coords) <- unique(dimnames(GPA$coords)[[3]])
ind.coords <- arrayspecs(ind.coords, p=ncol(ind.coords)/3, k=3)
# centroid size data
ind.Csize <- as.vector(aggregate(GPA$Csize ~ dimnames(GPA$coords)[[3]], FUN=mean)[,-1])
names(ind.Csize) <- unique(dimnames(GPA$coords)[[3]])
# ind.Csize - These are the individuals Csize data for plotting specimen allometry graphs

GPA_mean <-geomorph.data.frame(coords=ind.coords, Csize=ind.Csize)

```
#Read in classifier file - 

```{r}

write.csv(names(ind.Csize), file="../Data/Raw/Species_list.csv", row.names=FALSE, col.names=FALSE)


```
# To understand WHAT, running a PCA. To understand WHAT, computing the PCA eigenvalues.
### YOU HAVE TO MAKE SURE YOU ARE USING THE IND.COORDS NOT THE GPA COORDS HERE - WILL LET YOU DO THAT
```{r}
#Run PCA
PCA <- plotTangentSpace(GPA_mean$coords, label=dimnames(GPA_mean$coords)[[3]])

#PCA eigenvalues
PCA$pc.summary$importance

#If there is an obvious problem specimen, find out what number specimen it is on the GPA
dimnames(GPA$coords)[[3]]

plotRefToTarget(GPA$coords[,,6], GPA$coords[,,16], method="vector")

```

# Visualize main differences OF WHAT AND WHY? YOU MIGHT ALSO ADD SOME DESCRIPTION OF THE RESULTS HERE, WHICH YOU CAN LATER PUT INTO RESULTS SECTION.
```{r}

#Below is a starting point for customizing your PlotRefToTarget plots. I recommend playing with this! This then gets fed into plotRefToTarget as per below.
PRTT=gridPar(pt.bg="green", pt.size = 1)

plotRefToTarget(PCA$pc.shapes$PC1min, PCA$pc.shapes$PC1max, method="vector", mag=0.6,gridPars = PRTT)
plotRefToTarget(PCA$pc.shapes$PC2min, PCA$pc.shapes$PC2max, method="vector") #Back to standard grey
```


#TO UNDERSTAND/TEST...., TESTING FOR ALLLOMETRY (WHAT ABOUT ALLOMETRY ARE YOU TESTING?)

#Figure out allometry - try to do this yourself, check the geomorph guide here. Is there significant allometry? Can you plot it? Is there significant correlation betwen PC1, 2 or 3 and Allometry? As a proxy for size, use GPA$Csize.Try this, I can help you later tonight if you get stuck.

```{r}
MGD<-geomorph.data.frame(GPA_mean)
attributes(MGD) #Should be "coords" and "csize"

MarAllo<-procD.allometry(GPA_mean$coords~Csize, f2=NULL, f3=NULL, logsz=TRUE, data=MGD, iter=149, print.progress=FALSE)
summary(MarAllo)

```

#Try to get a stable tree code to work - find the phylogeny, try to read it in. I have some model code for you but try to use google and help files to see if you can read it in.

```{r}
#All 3 analyses (Maximum likelihood (ML), Bayesian, and parsimony) performed in Mitchell et al. (2014) resulted in well-resolved and concordant trees.
#I chose the first tree (ML, without outgroups) due to the use of it in their further analyses

tree=read.nexus("../Data/Raw/Mitchell_pruned_tree.nex")
tree_use <- tree[1]$tree_1

###~~~~~~~CHANGING SPECIES NAME~~~~~~~~~~###
tree_use$tip.label
tree_use$tip.label[166]  
#This shows particular tip label (without the number present this code will provide you with all the tip labels)

names(tree_use$tip.label[NUMBER] <- "New_Species")

names(tree_use$tip.label[5]) <- "Caluromys_derbianus"
names(tree_use$tip.label[1]) <- "Lestoros_inca"
names(tree_use$tip.label[5]) <- "Dendrolagus_inustus"
names(tree_use$tip.label[5]) <- "Philander_andersoni"
names(tree_use$tip.label[5]) <- "Lasiorhinus_krefftii"
names(tree_use$tip.label[5]) <- "Phalanger_mimicus"
names(tree_use$tip.label[5]) <- "Planigale_ingrami"
names(tree_use$tip.label[5]) <- "Sminthopsis_murina"
names(tree_use$tip.label[5]) <- "Sminthopsis_macroura"
names(tree_use$tip.label[5]) <- "Thylamys_elegans"

#to do the namecheck, Csize is easier to navigate to but it is part of the same gdf so that's fine
namecheck=name.check(tree_use,GPA_mean$Csize)

Matchtest=match(tree_use$tip.label,GPA_mean$coords)
length(which(!is.na (Matchtest)))


tree_synch=drop.tip(tree_use,namecheck$tree_not_data)

name.check(tree_synch)

```

#Create two partitions partitions
```{r}
### Import the  partition map
part.gp=as.matrix(read.csv("Data/Raw/partitions.csv", header=FALSE))

```

#Compare evolutionary rates of partitiosn
```{r}


```

