---
title: "MarsPCAGraphs"
output: html_document
---

```{r}
library(png)

setwd("D:/Github_Projects/Mars-shape/Data/")



```



```{r }
cols <- c("blue", "dark green", "#00CCCC", "#CC0000", "#0099FF", "light green",
"#9966FF", "#FF6600", "#FF6600", "#00CC00", "#FFCC00", "dark blue")
# specific colours for clades, use hex values or real names
names(cols) <- levels(speclist$Clade) # assign levels to colours
col.gp <- cols[match(speclist$Clade, names(cols))] # magic!
head(col.gp) # see the result
```

## Here we make a short form vector of species names of type “G.sp” for plotting:
```{r }

head(speclist$Specimen) # viewing the species names

Gen.sp <- matrix(unlist(strsplit(as.character(speclist$Specimen), "_")),
ncol=2, byrow = T)
# string split is used to separate a character string
head(Gen.sp)# should show a genus and species names split

Gen.sp <- paste(substr(Gen.sp[,1], 1, 1), substr(Gen.sp[,2], 1, 3), sep=".") 
head(Gen.sp) # see the result

```

##Preparing to make Shape Change Graphs (in 3D)

#Here we are going to warp a mesh to the shapes defined by the minima and maxima of the PCA
```{r}
ref <- mshape(GPA_Allmean$coords) # First get the mean shape
# Load in the PLY file and prepare for plotting
Dviv <- read.ply("Dasyurus_viverrinus_ascii")
Dviv$material <- "#31a354" ; open3d(); shade3d(Dviv) # change colour and view the mesh
mesh.coords <- read.morphologika("D:/Github_Projects/Mars-shape/Data/Raw/Coordinates/Dasyurus_viverrinus_J7996.txt")

# read in the landmarks digitized on this ply file in Checkpoint
plotspec(Dviv,mesh.coords[,,1]/1000, centered = F)
# check that the landmarks match the mesh; dividing by 1000 takes
# into account the scaling discrepancy of Checkpoint output files

refmesh <- warpRefMesh(Dviv, mesh.coords[,,1]/1000, ref, centered=FALSE)
# creates a mean shaped mesh

```


## preparing axis labels

```{r}

pc1lab <-paste("PC 1 ","(",
round(100*PCA_mean$pc.summary$importance[2,1],1),"%)",sep="")
pc2lab <-paste("PC 2 ","(",
round(100*PCA_mean$pc.summary$importance[2,2],1),"%)",sep="")
pc3lab <-paste("PC 3 ","(",
round(100*PCA_mean$pc.summary$importance[2,3],1),"%)",sep="")

pc1lab  #Shows you what it should look like

```



##Now to make a fancy PC plot - THIS IS DONE AS ONE BIG CHUNK

#plotting PCs 1-2, coloured by clade and particular clades as shaded convex hulls

```{r pressure, echo=FALSE}

png("..Data/Figure 3.png",width = 500, height = 900, units = "px", pointsize = 12)
# setting up to save the picture to file with those dimensions

mat <- matrix(c(1,1,2,2,3:8), nrow=5,byrow=T)
layout(mat, widths=c(1,1,0.5,0.5,0.5), heights=c(0.4,0.55,0.3,0.3,0.3))
# set the size of the rows and columns

#Plot 1 - PC1 vs PC3
par(mar=c(0, 4, 1, 1)) # sets the margins
plot(PCA_mean$pc.scores[,1], PCA_mean$pc.scores[,3], asp=T,xlab= "",
ylab= pc3lab, pch = 21, cex = 1.5, bg = col.gp, cex.axis=1.3, cex.lab=1.3,
bty="n", xaxt="n")
axis(side = 1, at=c(-0.2,0.1), labels = FALSE, col.ticks = "white")

# Add coloured polygons for the following clades:
clades <- levels(speclist$Clade)[c(1,2,4,7,11)]
for(i in clades){
  tmp <- which(speclist$Clade == i)
  EP <- rownames(PCA_mean$pc.scores[tmp,])[chull(PCA_mean$pc.scores[tmp, c(1,3)])] # find outer edge points
polygon(PCA_mean$pc.scores[EP,c(1,3)], col=adjustcolor(cols[which(names(cols) == i)],
  alpha.f = 0.3), border=cols[which(names(cols) == i)]) # Plot convex hull as polygon
}

# Then add the "Possums"
poss <- c(which(speclist$Clade == "Petauroid"),
which(speclist$Clade == "Phalangerid"))
EP <- rownames(PCA_mean$pc.scores[poss,])[chull(PCA_mean$pc.scores[poss, c(1,3)])] # find outer edge points
polygon(PCA_mean$pc.scores[EP,c(1,3)], col=adjustcolor(cols[which(names(cols) ==
"Phalangerid")], alpha.f = 0.3), border=cols[which(names(cols) == "Phalangerid")])

#Plot 2 - PC1 vs PC2
par(mar=c(4, 4, 0, 1)) # sets the margins
plot(PCA_mean$pc.scores[,1], PCA_mean$pc.scores[,2], asp=T, xlab= pc1lab, cex.axis=1.3, cex.lab=1.3,
ylab= pc2lab, pch = 21, cex = 1.5, bg = col.gp, bty="n") # plots PCA

# Add coloured polygons for the following clades:
clades <- levels(speclist$Clade)[c(1,2,4,7,11)]
for(i in clades){
tmp <- which(speclist$Clade == i & speclist$Locomotion == "no")
EP <- rownames(PCA_mean$pc.scores[tmp,])[chull(PCA_mean$pc.scores[tmp, c(1,2)])] # find outer edge points
polygon(PCA_mean$pc.scores[EP,c(1,2)], col=adjustcolor(cols[which(names(cols) == i)],
alpha.f = 0.3), border=cols[which(names(cols) == i)]) # Plot convex hull as polygon
}

#the "Possums"
EP <- rownames(PCA_mean$pc.scores[poss,])[chull(PCA_mean$pc.scores[poss, c(1,2)])] # find outer edge points
polygon(PCA$pc.scores[EP,c(1,2)], col=adjustcolor(cols[which(names(cols) ==
"Phalangerid")], alpha.f = 0.3), border=cols[which(names(cols) == "Phalangerid")])

remove(EP, tmp) # clean up environment

# Plot the 3D model screenshots below
par(mar=c(0, 0, 0, 0)) # sets the margins
for(i in 1:3){
tmp <- readPNG(paste("Figures/Mesh_PC",i,"min_lateral.png",sep=""))
plot(c(0,dim(tmp)[2]),c(0,dim(tmp)[1]),type="n", axes=F, xlab = "", ylab = "",asp=T)
rasterImage(tmp,0,0,dim(tmp)[2],dim(tmp)[1])
text(60,10, labels=paste("PC",i," min.", sep=""),cex=1.3)
tmp <- readPNG(paste("Figures/Mesh_PC",i,"max_lateral.png",sep=""))
plot(c(0,dim(tmp)[2]),c(0,dim(tmp)[1]),type="n", axes=F, xlab = "", ylab = "",asp=T)
rasterImage(tmp,0,0,dim(tmp)[2],dim(tmp)[1])
text(60,10, labels=paste("PC",i," max.",sep=""),cex=1.3)
}
dev.off() # end writing png file
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
